Towards a standardized evaluation of imputation methodology

### 2022-03-24 Formatted/Formatted.tex
Re-order missingness problems section

\begin{document}
%\DOIsuffix{bimj.DOIsuffix}
\DOIsuffix{bimj.200100000}
\Volume{52}
\Issue{61}
\Year{2010}
\pagespan{1}{}
\keywords{Evaluation; Imputation; Missing data; Simulation studies;\\[1pc]
\noindent \hspace*{-4.2pc} Supporting Information for this article is available from Github, \break \hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}.
}  %%% semicolon and fullpoint added here for keyword style

@@

The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology. These concerns stem from careful consideration with fellow 'imputers' and from encounters as a reviewer for statistical journals. Second, to provide imputation methodologists with a suggested course of action when simulating missing data problems. This suggested approach should identify a common ground, but is in no way intended as an absolute solution. This identifies the third purpose of this paper: discussion. We hope to elicit critical thinking with respect to the problems at hand. We are all convinced that our methodology has some merit. But for sake of progress it would be much more advantageous if the aim of our evaluations would go beyond \emph{proving the point} and would legitimately consider the statistical properties.

@@

An obvious problem with design-based simulation is that obtaining a large dataset without missing entries is very challenging. The idea behind this data generating mechanisim is to resample repeatedly from an empirical dataset which presents certain relevant characteristics but has an unknown true data generating mechanism. Since most real-world data contains at least one missing entry, and the non-response model underlying this missingness is by definition unknown, the simulator needs to deal with missingness in \emph{some} way before empirical data can serve as comparative truth in the simulations. It may seem as an intuitive solution to only draw complete cases from the incomplete dataset, which would indeed yield complete samples. However, there may be inherent differences between cases with and cases without any missing values in the incomplete dataset, due to the unknown non-response model. Only sampling complete cases would thus result in a simulation which fails to capture all relevant real-world conditions [which refutes the main reason for using a design-based data generating mechanism in the first place]. Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset from which samples may be drawn. Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.

@@

Something else to consider is that if missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly (whether it may be from a known model or empirical data). [Add pooling when the sample is the population citep{vink14} and write something about bootstrapping SEs and that SEs may be incorrect if the (re)sampling is not handled correctly.]  One needs to realize that the conventional pooling rules \citep[cf.][p. 76-77]{rubi87} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}. [in our own experience it may be difficult to get true parameter estimates of transformed data (e.g., the mean of a variable with skewness).] % we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar. we verplaatsen liever de monte carlo naar het niveau van de missingness.

@@

Missingness comes in many shapes and forms [platitude, but how else to start?]. The most obvious characteristics of missing data are the missingness pattern(s) and missingness mechanism(s). Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data \citep[][p. 8]{litt20}.

@@

The missing data pattern may inadvertently guide which imputation method prevails in simulation studies. Some ad hoc methods are known to yield valid inferences in very specific missing data patterns (e.g., list-wise deletion when there is missingness in the outcome variable of the analysis model exclusively), while they bias inferences in any other case. Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions [re-read this! and add monotone/univariate vs multivariate and sporadic vs systematic dichotomies]. Related to the missingness pattern across variables is the missingness proportion. The terminology on this topic can be confusing. Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry? We will refer to the latter as the proportion of incomplete cases. But this is not a clear distinction in the literature, which may lead to incorrect conclusions [re-phrase this too]. Moreover, the amount of missingness that is used to simulate the performance of imputation methods differs between studies. [Add ref to chapter \citep{liu21} and add info.] Some studies use only 10\% missing data where other studies push the limits to additionally investigate the performance under at least 50\% missing data. This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions. This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.

[Briefly introduce missingness mechanisms here and show potential problems, but without simulation results (moved to separate section). This part should contain: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \emph{spuriously} if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice. And finally something about types of M(N)AR, but what are specific problems with that? ref dance paper rianne]

@@

With MCAR missingness mechanisms, the probability to be missing is the same for all cases. This is a necessary simulation condition for evaluating the performance of imputation procedures. If an imputation method is not able to solve the problem (i.e. yield valid inference) under MCAR, the statistical properties of the procedure are not sound. Sadly, the straightforward case of MCAR is often neglected from simulation studies and focus is drawn to the evaluation of MAR mechanisms only. Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in my view may be far too simplistic.

@@

When MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid. When MAR is induced from data without multivariate relations, the inferential implication of the missingness would mimic those of MCAR. This problem is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications. Figure \ref{fig:MAR} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.

\begin{figure}[htb]
\begin{center}
\resizebox{\textwidth}{!}{
\includegraphics[]{figures/spurious_MAR.pdf}}
\caption{Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$). Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$.}
\end{center}
\label{fig:MAR}
\end{figure}
\vspace*{1pc}

It is obvious that the MAR missingness generated from data with low correlation would yield similar inference to the MCAR missingness, even though the missingness is random. In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p. 48]{fimd}: the missingness does not depend on the incomplete variable. This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.

@@

\subsection{Performance evaluation problems}

[Add bridge from last section. And explain how this part is subdivided: first general characteristics of the imputations such as convergence, then performance measures.]

[Performance of different imputation methods may be influenced by preconditions of the methods themselves. All imputation models should be congenial, but this is difficult to test to often ignored even in single imputation endeavor, let alone in simulation studies with 100s of imputations. Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies. Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others.]

@@

[Then onto the performance measure part.]

The evaluation criteria used to assess imputation performance vary from one developer to another. This is not surprising as people from different fields could have a different focus on the problem at hand. But the choice of performance measure(s) may inadvertently distort statistical properties of the imputed data.

Developers often only inquire about the 'accuracy' (i.e. how well can the method reproduce the original data). [Add RMSE at cell level here?] The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. We are interested in the correct answer to the research question; not in the truth itself. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied. After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p. 591]{neym34}. [So, what could go wrong by focusing on accuracy exclusively?]

@@

@@

Sampling variation has always been an essential part of the evaluation of multiple imputation methodology. However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account \citep{vink14}. After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.

@@

\begin{table}[htb]
\begin{center}
\caption{Steps to consider in imputation simulation studies.}
\begin{tabular}{lll}
\hline
Step                  & Aspects under consideration \\
\hline
1. Set scope          & Aim(s), missing data method(s) to evaluate, number of simulation \\
                      & repetitions \\
2. Obtain truth       & Data generating mechanism(s), estimand(s), sampling variance \\
3. Induce missingness & Missingness mechanism(s), missing data pattern(s), missingness \\
                      & proportion(s) \\
4. Apply methods      & Imputation model(s), analysis model(s) \\
% Benchmarking, imputation model parameter, estimand
5. Evaluate methods   & Imputation model fit, performance measures \\
6. Report             & Text, visualization, checklist, simulation script \\
\hline
\end{tabular}
\end{center}
% \footnotesize{Note. Multiple aspects may be varied across simulation conditions (e.g., using a factorial design with several data generating mechanisms and missingness proportions).}
\end{table}

@@

Make aim clear. Choose scope (type of DGM, target) and imputation methods to evaluate. Ideally, perform fully factorial design, see \citet{morr18}. Think about tolerable level of uncertainty to pick a number of simulation repetitions. Think about general set-up: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?

@@

[Decide which data genetation mechanism to use and obtain true value of the estimand(s).] In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations. In design-based simulation, data are sampled from a sufficiently large set, so the parameters of this specific set serve as the comparative truth in the simulations.

@@

First, one should always consider MCAR missingness, i.e. the scenario where the missing values are missing at random and the observed values are observed at random. Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.

@@

Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness. Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model. This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

Fourth, the amount of missingness must be varied. Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process. Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.

The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness. If there is no missingness - or if there is no data, for that matter - the inference does not depend on the assumption. Alternatively, the validity of assumptions become increasingly important when the missingness increases. Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions. I therefor propose, for all mechanisms, to evaluate at least the following univariate missingness percentages when evaluating imputation routines:

\begin{itemize}
  \item 10\% missingness: Depending on the size of the data, this percentage can be considered as a lower bound of realistic evaluation. Anything less than 10\% may be of little influence on the true data inference. Performance of a missing data method should at least be acceptable for most missing data problems. 
  \item 25\% missingness: This is a fair amount of missingness and will, depending on the observed data information, have a noticeable influence on the completed data inference. When compared to the condition with 10\% missingness, the inference obtained under 25\% missingness should be less certain (i.e. confidence/credibility interval width should increase), but estimates should still be properly covered and the statistical properties of the missing data method should be sound. In practice, at least to my experience in social sciences and official statistics, 25\% univariate missingness can easily be considered as a realistic missingness percentage. 
  \item 50\% missingness: Performance under 50\% percent simulated missingness will most likely be impacted severely. Depending on how the missingness mechanism interacts with the simulated data, some imputation techniques may yield estimates that are under-covered such that the completed data inference should not be deemed valid anymore. If a method yields acceptable inference under 50\% MAR missingness, we can determine that the statistical properties of the imputation methodology are sound. 
\end{itemize}
Although I limit the focus here to ignorable non-response, the above suggested proportions are equally applicable to simulations under non-ignorable non-response.

Fifth, the used missingness mechanism should be detailed, either graphically or written as a function of the data. Often, when inducing missingness, authors remain vague about the actual missingness mechanism under investigation and, even worse, some authors only report something like
\begin{quote}
\emph{We generated missing data following a MAR missingness mechanism.}
\end{quote}
This should be considered unacceptable as claims about the validity of the multiple imputation inference depends heavily on the simulated missingness mechanism.

@@

@@

\subsection{Step 4. Evaluate performance}

% 4a. Evaluate imputations: check failed methods, algorithmic convergence (if applicable)
% 4b. Check imputation model fit: anomalies in distributionsal characteristics, plausibility of imputed values (if requested), PPC
% 4c. Apply performance measures: at least consider bias, coverage, CI length, if applicable also RMSE (at prediction and/or cell level)
The evaluation criteria may depend on the estimand. When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored. Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice. In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
\begin{itemize}
  \item \emph{Bias:} Results should preferable be unbiased. However, the way bias is considered can greatly influence the interpretation of the results. For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered. The way bias is computed should therefore be carefully chosen and described. 
  \item \emph{Interval width:} The way the confidence interval is calculated should be described. Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference. However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore. 
  \item \emph{Coverage:} Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage of 95\% would be most efficient. Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained. 
  \item \emph{RMSE:} [RMSE at cell level and at prediction level (if applicable).]
  \item \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data. Under anything but the MCAR assumption, this can be expected. When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for an detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}. When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study. 
  \item \emph{Plausibility of the imputed values:} Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference. However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property. When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values. 
  \item \emph{Fit of the imputation model:} [PPC, ref Mingyang, \citep{nguy17, zhao22}.]
  \item \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}. The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.  
  \end{itemize}

@@

\subsection{Step 5. Report}

@@

>>>>

\begin{document}
%\DOIsuffix{bimj.DOIsuffix}
\DOIsuffix{bimj.200100000}
\Volume{52}
\Issue{61}
\Year{2010}
\pagespan{1}{}
\keywords{Evaluation; Imputation; Missing data; Simulation studies;\\ [1pc]
% \noindent \hspace*{-4.2pc} Supporting Information for this article is available from Github, \break \hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}.
% } %%% semicolon and fullpoint added here for keyword style
\noindent \hspace*{-4pc} Supporting Information for this article is available online from \\
\hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}. 
}  %%% semicolon and fullpoint added here for keyword style

@@

The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology. These concerns stem from careful consideration with fellow 'imputers' and from encounters as a reviewer for statistical journals. Second, to provide imputation methodologists with a suggested course of action when simulating missing data problems. This suggested approach should identify a common ground, but is in no way intended as an absolute solution. This identifies the third purpose of this paper: discussion. We hope to elicit critical thinking with respect to the problems at hand. We are all convinced that our methodology has some merit. But for sake of progress it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties.

@@

An obvious problem with design-based simulation is that obtaining a large dataset without missing entries is very challenging. The idea behind this data generating mechanisim is to resample repeatedly from an empirical dataset which presents certain relevant characteristics but has an unknown true data generating mechanism. Since most real-world data contains at least one missing entry, and the non-response model underlying this missingness is by definition unknown, the simulator needs to deal with missingness in \textit{some} way before empirical data can serve as comparative truth in the simulations. It may seem as an intuitive solution to only draw complete cases from the incomplete dataset, which would indeed yield complete samples. However, there may be inherent differences between cases with and cases without any missing values in the incomplete dataset, due to the unknown non-response model. Only sampling complete cases would thus result in a simulation which fails to capture all relevant real-world conditions [which refutes the main reason for using a design-based data generating mechanism in the first place]. Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset from which samples may be drawn. Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.

@@

Something else to consider is that if missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly (whether it may be from a known model or empirical data). Sampling variation has always been an essential part of the evaluation of multiple imputation methodology. However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account \citep{vink14}. After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies. Note that the conventional pooling rules \citep[cf.][p. 76-77]{rubi87} do not apply for finite population inference. Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}. [An example from our own experience is that it may be difficult to get true parameter estimates of transformed data (in this case, the mean of a variable with a skew normal distribution), If we wouldn't have used a simulation set-up without sampling variance, it may have led to imprecise reference values of the estimands in our simulation. So, on the one hand, the typical set-up obfuscates the process that we're interested in (i.e., varying the missing data problem over simulation repetitions), while on the other hand, SEs may be incorrect if the sampling variance is not handled correctly in subsequent inferences.] % we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar. we verplaatsen liever de monte carlo naar het niveau van de missingness.

@@

In order to evaluate the effectivity of imputation methodology to recover the truth established from a chosen data generating mechanism, some form of missingness has to be induced. Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like
\begin{quote}
\textit{We generated missing data following a missing at random missingness mechanism.}
\end{quote}
This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness patterns and missingness mechanisms. Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data \citep[][p. 8]{litt20}. Under this definition, the missingness pattern also encompasses the proportion of missingness in a given dataset.

@@

Even the terminology on missingness generation can be confusing. Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry? In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term 'missingness proportion' restricted to the variable-by-variable interpretation. This distinction, however, is not always clear in the existing literature. Convoluting the two concepts may lead to incorrect generalizations from simulation studies, since the two are rarely equal (e.g., a proportion of 50\% incomplete cases could translate to a missingness proportion of just 25\% in bivariate data, or even 10\% in data with five variables). Moreover, the value of the actual missingness proportion could be diffusing too. Some studies use only 10\% missing data where other studies push the limits to additionally investigate the performance under at least 50\% missing data. This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions. This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.

% Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process. Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data. The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness. If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption. Alternatively, the validity of assumptions become increasingly important when the missingness increases. Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.

Other problems with missingness patterns entail the distribution of missing values across cases (as opposed to variables). [Add: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivatiate patterns.] Specific missing data patterns may inadvertently guide which imputation method prevails in performance. Some ad hoc methods are known to yield valid inferences under certain restricted conditions (e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively), while they bias inferences in any other case. Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions.

[Missingness mechanisms: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice. And finally something about types of M(N)AR, but what are specific problems with that? ref dance paper rianne]

% MCAR missingness: The scenario where the missing values are missing at random and the observed values are observed at random. Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.

% Missingness type: Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

%  the validity of assumptions become increasingly important when the missingness increases.

@@

With MCAR missingness mechanisms, the probability to be missing is the same for all cases. This is a necessary simulation condition for evaluating the performance of imputation procedures. If an imputation method is not able to solve the problem (i.e. yield valid inference) under MCAR, the statistical properties of the procedure are not sound. Sadly, the straightforward case of MCAR is often neglected from simulation studies and focus is drawn to the evaluation of MAR mechanisms only. Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in our view may be far too simplistic.

@@

When MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid. When MAR is induced from data without multivariate relations, the inferential implication of the missingness would mimic those of MCAR. This problem is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications.

% Figure \ref{fig:MAR} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.

% \begin{figure}[htb]
% \begin{center}
% \resizebox{\textwidth}{!}{
% \includegraphics[]{figures/spurious_MAR.pdf}}
% \caption{Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$). Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$.}
% \end{center}
% \label{fig:MAR}
% \end{figure}
% \vspace*{1pc}

A MAR missingness generated from data with low correlation would yield similar inference to the MCAR missingness, even though the missingness is random. In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p. 48]{fimd}: the missingness does not depend on the incomplete variable. This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.

@@

\subsection{Evaluation problems}

% The evaluation criteria may depend on the estimand. When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored. Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.

% For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered.

[Add bridge from last section. And explain how this part is subdivided: first problems with evaluation of imputations (general characteristics of the imputations such as convergence and plausibility), then evaluation of performance (in terms of actual performance measures).]

[All imputation models should be congenial, but this is difficult to test to often ignored even in single imputation endeavor, let alone in simulation studies with hundreds of imputations. Checks of imputation model fit are often omitted. Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others. Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies. Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]

@@

[Add segway between evaluation of imputation and actual performance measures.]
The evaluation criteria used to assess imputation performance vary from one developer to another. This is not surprising as people from different fields could have a different focus on the problem at hand. But the choice of performance measure(s) may inadvertently distort statistical properties of the imputed data.

@@

Developers often only inquire about the 'accuracy' (i.e. how well can the method reproduce the original data). [Add RMSE at cell level here?] The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. We are interested in the correct answer to the research question; not in the truth itself. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied. After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p. 591]{neym34}. [So, what could go wrong by focusing on accuracy exclusively? See problems with RMSE of predictions.]

@@

@@

\begin{table}[htb]
\begin{center}
\caption{Steps to consider in imputation simulation studies.}
\begin{tabular}{lll}
\hline
Step                    & Aspects under consideration \\
\hline  
1. Set scope            & Aim(s), missing data method(s) to evaluate, number of simulation \\
                        & repetitions \\
2. Obtain truth         & Data generating mechanism(s), estimand(s), sampling variance \\
3. Induce missingness   & Missingness mechanism(s), missing data pattern(s), missingness \\
                        & proportion(s) \\
4. Apply methods        & Imputation model(s), analysis model(s) \\
5. Evaluate imputations & Algorithmic convergence, distributional characteristics, imputation \\
                        & model fit \\
6. Evaluate performance & Performance measures,  \\
7. Report               & Text, visualization, checklist, simulation script \\
\hline
\end{tabular}
\end{center}
% \footnotesize{Note. Multiple aspects may be varied across simulation conditions (e.g., using a factorial design with several data generating mechanisms and missingness proportions).}
\end{table}

@@

[Make aim clear. Choose scope (type of DGM, target) and imputation methods to evaluate. Ideally, perform fully factorial design, see \citet{morr18}. Think about tolerable level of uncertainty to pick a number of simulation repetitions. Think about general set-up: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?]

@@

[Decide which data genetation mechanism to use and obtain true value of the estimand(s).]

In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations. In design-based simulation, data are sampled from a sufficiently large set, so the parameters of this specific set serve as the comparative truth in the simulations.

@@

[Missingness mechanisms]
First, one should always consider MCAR missingness, i.e. the scenario where the missing values are missing at random and the observed values are observed at random. Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.

@@

[Missingness type]
Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness. Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

[Missingness proportion]
Fourth, the amount of missingness must be varied. Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process. Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data. The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness. If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption. Alternatively, the validity of assumptions become increasingly important when the missingness increases. Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions. We therefore propose, for all mechanisms, to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness see Table 2.

\begin{table}[htb]
\begin{center}
\caption{Suggested univariate missingness proportions.}
\begin{tabular}{lll}
\hline
Prop. & Reasoning for suggested missingness proportion \\
\hline  
10\%  & Depending on the size of the data, this percentage can be considered as a lower \\
      & bound of realistic evaluation. Anything less than 10\% may be of little influence on \\
      & the true data inference. Performance of a missing data method should at least be \\
      & acceptable for most missing data problems. \\ & \\
25\%  & This is a fair amount of missingness and will, depending on the observed data \\
      & information, have a noticeable influence on the completed data inference. When \\
      & compared to the condition with 10\% missingness, the inference obtained under \\
      & 25\% missingness should be less certain (i.e. confidence/credibility interval width \\
      & should increase), but estimates should still be properly covered and the statistical \\
      & properties of the missing data method should be sound. In practice, at least to our \\
      & at least to my experience in social sciences and official statistics, 25\% \\
      & univariate missingness can easily be considered as a realistic missingness percentage. \\ & \\
50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
      & severely. Depending on how the missingness mechanism interacts with the simulated \\
      & data, some imputation techniques may yield estimates that are under-covered such that \\
      & the completed data inference should not be deemed valid anymore. If a method yields \\
      & acceptable inference under 50\% MAR missingness, we can determine that the statistical \\
      & properties of the imputation methodology are sound. \\
\hline
\end{tabular}
\end{center}
\end{table}

Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.

@@

[Use CCA as benchmark method: the imputation method should outperform CCA. Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matix, regularization if applicable, etc.). Estimate estimand(s) using the analysis model, use correct pooling rules.]

@@

\subsection{Evaluate imputations}

% 4a. Evaluate imputations: check failed methods, algorithmic convergence (if applicable)
% 4b. Check imputation model fit: anomalies in distributionsal characteristics, plausibility of imputed values (if requested), PPC

Convergence of the algorithm: Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}. The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.

Distributional characteristics: In practice, the distribution of the incomplete data may differ greatly from the observed data. Under anything but the MCAR assumption, this can be expected. When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for an detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}. When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.

Fit of the imputation model: [PPC, ref Mingyang, \citep{nguy17, zhao22}.]

Plausibility of the imputed values: Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference. However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property. When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.

@@

\subsection{Evaluate performance}

% 4c. Apply performance measures: at least consider bias, coverage, CI length, if applicable also RMSE (at prediction and/or cell level)

The evaluation criteria may depend on the estimand. When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored. Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.

In general, we would say that each multiple imputation routine should be evaluated on at least the following points:

Bias: Results should preferable be unbiased. However, the way bias is considered can greatly influence the interpretation of the results. For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered. The way bias is computed should therefore be carefully chosen and described.

Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage of 95\% would be most efficient. Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.

Interval width: The way the confidence interval is calculated should be described. Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference. However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.

RMSE: [RMSE at cell level and at prediction level (if applicable).]

\subsection{Report}

@@

The used missingness mechanism should be detailed, either graphically or written as a function of the data.
